Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: preliminary report

### 2022-04-25 protocol/protocol.Rmd

---
title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: protocol"
output:
  html_document:
    self_contained: yes
  pdf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
  rtf_document:
    toc: no
  word_document: default
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

@@

The following primary outcomes were considered: 
* Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. 
* COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc. This outcome also includes COVID-19 admissions (see below) to pick up more severe disease episodes that may have bypassed A&E. 
* COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes. 
* COVID-19 death. Deaths were classified as from COVID-19 if deaths with the U07.1 or U07.2 ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
* Non-COVID-19 death.

@@

Essential matching criteria:
* date of booster dose
* primary course vaccine brand
* date of second vaccine dose, within 7 days
* evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission)
* clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither)
* age, within 3 years, and age groups used by JCVI for prioritisation
* Geographical region, one of the below in decreasing order of precision:
  * Lower Layer Super Output Area (LSOA)
  * Middle Layer Super Output Area (MSOA)
  * NHS Strategic Health Partnership (STP)
  * NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West)

Additional matching criteria if possible:
* English Index of Multiple Deprivation (IMD), within X days (tbd)
* ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census);
* existing morbidity count: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease
* learning disabilities
* severe mental illness
* the number of SARS-CoV-2 tests in the 6 months prior to boosting
* elective in-hospital episode at the time of boosting

@@

Confidence limits for the KM estimate will be derived from complimentary log-log transformation.
Confidence limits for the risk difference will be derived from the square root of the sum of squares the treatment-specific KM standard errors. The KM standard errors will use Greenwood's formula.
Confidence limits for the relative risk??

Alternatively, we could boostrap.

@@

# Results

## Study population and matching

`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="moderna") %>% pull(pct_all))`) eligible for matching. `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).

#### Table 1: inclusion criteria

```{r table-flowchart, echo=FALSE, warning=FALSE, message=FALSE}

match_flowchart %>%
  select(-crit) %>%
   pivot_wider(
    id_cols = c(criteria, vax3_type),
    names_from = vax3_type,
    values_from = c(n, n_exclude, pct_all)
  ) %>%
  gt() %>%
  cols_label(
    criteria = "",

    n_pfizer = "N",
    n_moderna = "N",

    n_exclude_pfizer = "N excluded",
    n_exclude_moderna = "N excluded",

    pct_all_pfizer = "% remaining",
    pct_all_moderna = "% remaining",
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("pfizer")
  ) %>%
   tab_spanner(
    label = "Moderna mRNA-1273",
    columns = ends_with("moderna")
  ) %>%
  # tab_spanner(
  #   label = "mRNA-1273",
  #   columns = ends_with("moderna")
  # ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  )
```

\newpage 
#### Figure 1: Cumulative matching success

```{r, figure-matchingsuccess, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(match_coverage$vax3_date )
xmax <- max(match_coverage$vax3_date )+1

plot_coverage_cumuln <-
  match_coverage %>%
  mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
    status_descr = fct_recoderelevel(as.character(status), recoder$status),
    cumuln = cumuln*((treatment*2) - 1)
  ) %>%
  ggplot()+
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=paste0(treatment,status_descr),
      fill=treatment_descr,
      alpha=status_descr,
      colour=NULL
    ),
    position=position_stack(reverse=TRUE),
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
    limits = c(xmin-1, NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  scale_alpha_discrete(range= c(0.8,0.4))+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom"
  )+
  NULL

plot_coverage_cumuln
cat("  \n\n")
```

#### Table 2: baseline characteristics

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
table1 <- 
  match_table1 %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(match_table1by$by_col, match_table1by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

## Estimated comparative effectiveness

### Main analysis

#### Table 3: Kaplan-Meier cumulative incidence estimates

```{r table-km, echo=FALSE, warning=FALSE, message=FALSE}

km_contrasts_overall %>%
  transmute(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,
    n.atrisk=n.atrisk_0,
    cml.atrisk_0=label_number(1)(cml.atrisk_0/7), 
    cml.atrisk_1=label_number(1)(cml.atrisk_1/7),
    n.event_0, 
    n.event_1,
    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
    kmrr = paste0(label_number(0.01)(kmrr)),
  ) %>%
  gt(groupname_col="outcome_descr") %>%
  cols_label(

    outcome_descr = "Outcome",
    subgroup_descr = "Subgroup",
    subgroup_level_descr = "Subgroup",

    n.atrisk = "N per group",

    cml.atrisk_0 = "Person-weeks", 
    cml.atrisk_1 = "Person-weeks",

    n.event_0 = "Events",
    n.event_1 = "Events",

    riskCI_0 = "KM risk (95% CI)",
    riskCI_1 = "KM risk (95% CI)",

    kmrdCI = "Risk-difference (95% CI)",
    kmrr = "Relative-risk"
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Moderna",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison",
    columns = starts_with("km")
  )

#### Figure 3: Kaplan-Meier cumulative incidence estimates

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=time, y=1-surv))+
  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  facet_grid(rows=vars(outcome_descr))+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y="Cumulative incidence",
    colour=NULL,
    title=NULL
  )+
  theme_minimal()+
  theme(
    axis.line.x = element_line(colour = "black"),
    panel.grid.minor.x = element_blank(),
    legend.position=c(.05,.95),
    legend.justification = c(0,1),
  )

plot_km
```

@@

## References

>>>>

---
title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: protocol"
output:
  word_document: default
  pdf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
  rtf_document:
    toc: no
  html_document:
    self_contained: yes
#bibliography: references.bib
#csl: vancouver.csl
#link-citations: yes
#zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

@@

The following primary outcomes were considered: 
* Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. 
* COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes. This outcome also includes COVID-19 admissions (see below) to pick up more severe disease episodes that may have bypassed A&E. 
* COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes. 
* COVID-19 death. Deaths were classified as from COVID-19 if deaths with the U07.1 or U07.2 ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
* Non-COVID-19 death.

@@

Essential matching criteria:

* date of booster dose
* primary course vaccine brand
* date of second vaccine dose, within 7 days
* evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission)
* clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither)
* age, within 3 years, and age groups used by JCVI for prioritisation
* Geographical region, one of the below in decreasing order of precision:

  * Lower Layer Super Output Area (LSOA)
  * Middle Layer Super Output Area (MSOA)
  * NHS Strategic Health Partnership (STP)
  * NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West)

Additional matching criteria if possible:

* English Index of Multiple Deprivation (IMD), within X days (tbd)
* ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census);
* existing morbidity count: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease
* learning disabilities
* severe mental illness
* the number of SARS-CoV-2 tests in the 6 months prior to boosting
* elective in-hospital episode at the time of boosting

@@

Confidence limits for the KM estimate will be derived from complimentary log-log transformation.
Confidence limits for the risk difference will be derived from the square root of the sum of squares the treatment-specific KM standard errors. The KM standard errors will use Greenwood's formula.
Confidence limits for the survival ratio are straight-forward, and can be based on the standard error of the cumulative hazard. But confidence limits for the risk ratio (or relative risk) are more complicated. TBD.

Alternatively, we could bootstrap.

@@

@@

### 2022-04-25 write-up/report.Rmd

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=time, y=1-surv))+
  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  facet_grid(rows=vars(outcome_descr))+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y="Cumulative incidence",
    colour=NULL,
    title=NULL
  )+
  theme_minimal()+
  theme(
    axis.line.x = element_line(colour = "black"),
    panel.grid.minor.x = element_blank(),
    legend.position=c(.05,.95),
    legend.justification = c(0,1),
  )

plot_km
```

>>>>

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=period_end, y=1-surv))+
  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  facet_grid(rows=vars(outcome_descr), scales="free_y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y="Cumulative incidence",
    colour=NULL,
    title=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position=c(0.05,0.1),
      legend.justification = c(0,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),

      panel.spacing = unit(0.8, "lines"),

    )+
    NULL

plot_km

plot_kmrd <- km_contrasts_daily %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
  geom_step(aes(x=period_start, y=kmrd))+
  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
  facet_grid(rows=vars(outcome_descr), scales="free_y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y="Cumulative risk difference",
    colour=NULL,
    title=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),

      panel.spacing = unit(0.8, "lines"),

    )+
    NULL

plot_kmrd

  plot_combined <- patchwork::wrap_plots(plot_km, plot_kmrd, plotkmrr, ncol=1)
  ggsave(
    filename="km.png",
    path=output_dir_rmd,
    plot=plot_combined,
  )

plot_combined
